
---
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

<style type="text/css">
.main-container {
  max-width: 90%;
  margin-left: auto;
  margin-right: auto;
  margin-bottom: 4em;
  display: flex;
  flex-direction: column;
  align-items: center;
}

.plotly{
  <!-- width: 100%; -->
  align: "center";
  align-items: center;

}

.svg-container{
  <!-- width: 100%; -->
  align: "center";
  margin: auto;
  position: relative;
}

.html-widget {
  margin: auto;
  min-width: 800px;
  <!-- margin-top: 30px; -->
}

.pp-box {
  display: flex;
  align-items: "center";
  flex-direction: column;
  margin-bottom: 50px;
}

.pp-header {
  margin-top: 25px;
}

<!-- .summary { -->
<!--   padding-left: 20px; -->
<!-- } -->

<!-- g.draglayer.cursor-crosshair { -->
<!--   width: 600px; -->
<!--   height: 600px; -->
<!-- } -->

p, #opener, #footer {
  max-width: 50vw;
  min-width: 600px;
  <!-- width: 500px; -->
  margin: auto;
  <!-- margin-left: auto; -->
  <!-- margin-right: auto; -->
  margin-bottom: 25px;

}

p {
  font-size: 12pt;
  display: relative;
}

#overall {
  display: flex;
  flex-direction: column;
  align-items: center;
  max-width: 650px;
}

#opener {
  <!-- max-width: 50vw; -->
  <!-- min-width: 500px; -->
  <!-- width: 500px; -->
  margin: auto;

}

.ani-plot-container {
  padding-right: 20px;
  margin-bottom: 25px;
}

#title {
  margin-top: 25px;
  width: 500px;
}

#byline {
  margin-bottom: 25px;
}

#footer {
  margin-top: 50px;
  align: "center";
}

#footer > p {
  font-size: 10pt;
}


</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align = 'center')
```

<div id = "overall">

<div id = "opener">

<h1 id = "title">Affective Trajectories</h1>
<h4 id = "byline">by Liz Cory</h4>

We all experience our feelings shifting — throughout the day and across periods of our lives. For some of us, the changes are slight, while others may feel totally different from one day to the next. To better understand how our feelings work and change, psychologists map them along two dimensions: "valence" and "arousal." Valence captures how pleasant you feel, and arousal marks how activated you feel. Together, valence and arousal are termed "affect."


This project documents how six individuals move through affective space over a two-week period, or, their <i>affective trajectories</i>. By animating daily reported feelings, we can explore where in affective space each person spends the most time across the sampling period. We can also see what emotion words a person uses to describe their feelings at each location by hovering over the dots.

</div>


---


```{r fonts, echo=FALSE,fig.show='hide'}
## Set up font
library(gfonts)

# setup_font(
#   id = "ibm-plex-sans",
#   output_dir = "./fonts"
# )

use_font("ibm-plex-sans", "fonts/css/ibm-plex-sans.css")
```

```{r libs}
library(tidyverse)
library(here)
library(readxl)
library(gganimate)
library(janitor)
library(ggiraph) 
library(plotly)
library(randomcoloR)
library(wesanderson)
```

```{r load data, include=FALSE}
## identify list of files we want
pp_files <- list.files(here("dat"), pattern="*.csv", full.names = TRUE)

## read in all those files
df.list <- lapply(pp_files, read.csv)
names(df.list) <- paste0("ppdf_", seq_len(6))

## add each df to the global environments so we can more easily manipulate in ggplot
list2env(df.list, envir=.GlobalEnv)

# get colors
colors <- c(wes_palette("Darjeeling1"), wes_palette("Rushmore1")[4])

```

<h3 class = "pp-header">Participant 1</h3>
<div class ="pp-box" align = "center">
<div class= "ani-plot-container">
```{r circumplex PP1}
#--------------------#
# add summary day of all points
# ppdf_1 <- rbind(ppdf_1,ppdf_1) %>% 
#   select(-c(X))
# 
# ppdf_1$sampling_day[(nrow(ppdf_1)/2 + 1):nrow(ppdf_1)] <- "99"
# 
# ppdf_1$sampling_day <- as.numeric(ppdf_1$sampling_day)

# ppdf_1$sampling_day <- factor(ppdf_1$sampling_day, levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,"sum"))

# plot all points
pd1p <- ggplot(ppdf_1, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = sampling_day),
             color = colors[1],
             alpha = 0.4,width = 5) +
  # ggtitle("Participant 1") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd1p2 <- ggplot(ppdf_1, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[1],
             alpha = 0.4,width = 5) +
  ggtitle("All sampling days") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "IBM Plex Sans",
  size = 14,
  color = toRGB("grey50"))

(ly1 <- ggplotly(pd1p, width = 610, height = 672) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  animation_opts(800))

# ggsave(here("fig",file = "pd1p.png"), width = 12, height = 12)

```
</div>

```{r 1over, fig.width=6,fig.height=6, fig.align='center', fig.show='hold'}

(ly1_2 <- ggplotly(pd1p2, width = 600, height = 600) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  config(displayModeBar = FALSE))

# pd1p + ggtitle("All sampling days")

```
</div>

<h3 class = "pp-header">Participant 2</h3>
<div class ="pp-box" align = "center">
<div class= "ani-plot-container">
```{r circ PP2, fig.width=6,fig.height=6, fig.align='center'}

# # add summary day of all points
# ppdf_2 <- rbind(ppdf_2,ppdf_2) %>% 
#   select(-c(X))
# 
# ppdf_2$sampling_day[(nrow(ppdf_2)/2 + 1):nrow(ppdf_2)] <- 99
# # ppdf_2$sampling_day <- factor(ppdf_2$sampling_day, levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,"sum"))

# plot all points
pd2p <- ggplot(ppdf_2, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = sampling_day),
             color = colors[2],
             alpha = 0.4,width = 5) +
  # ggtitle("Participant 2") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd2p2 <- ggplot(ppdf_2, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[2],
             alpha = 0.4,width = 5) +
  ggtitle("All sampling days") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "IBM Plex Sans",
  size = 14,
  color = toRGB("grey50"))

(ly2 <- ggplotly(pd2p, width = 610, height = 672) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  animation_opts(800))

# ggsave(here("fig",file = "pd2p.png"), width = 12, height = 12)


```
</div>

```{r 2over, fig.width=6,fig.height=6, fig.align='center', fig.show='hold'}

(ly2_2 <- ggplotly(pd2p2, width = 600, height = 600) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  config(displayModeBar = FALSE))

# pd2p + ggtitle("All sampling days")

```
</div>

<h3 class = "pp-header">Participant 3</h3>
<div class ="pp-box" align = "center">
<div class= "ani-plot-container">
```{r circ PP3, fig.width=6,fig.height=6, fig.align='center', fig.show='hold'}
 
# add summary day of all points
# ppdf_3 <- rbind(ppdf_3,ppdf_3) %>% 
#   select(-c(X))
# 
# ppdf_3$sampling_day[(nrow(ppdf_3)/2 + 1):nrow(ppdf_3)] <- 99
# # ppdf_3$sampling_day <- factor(ppdf_3$sampling_day, levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,"sum"))

# plot all points
pd3p <- ggplot(ppdf_3, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = sampling_day),
             color = colors[3],
             alpha = 0.4,width = 5) +
  # ggtitle("Participant 3") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd3p2 <- ggplot(ppdf_3, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[3],
             alpha = 0.4,width = 5) +
  ggtitle("All sampling days") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "IBM Plex Sans",
  size = 14,
  color = toRGB("grey50"))

(ly3 <- ggplotly(pd3p, width = 610, height = 672) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  animation_opts(800))

# ggsave(here("fig",file = "pd3p.png"), width = 12, height = 12)


```
</div>

```{r 3over, fig.width=6,fig.height=6, fig.align='center', fig.show='hold'}

(ly3_2 <- ggplotly(pd3p2, width = 600, height = 600) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  config(displayModeBar = FALSE))

# pd3p + ggtitle("All sampling days")

```
</div>

<h3 class = "pp-header">Participant 4</h3>
<div class ="pp-box" align = "center">
<div class= "ani-plot-container">
```{r circ PP4, fig.width=6,fig.height=6, fig.align='center', fig.show='hold'}

# add summary day of all points 
# ppdf_4 <- rbind(ppdf_4,ppdf_4) %>% 
#   select(-c(X))
# 
# ppdf_4$sampling_day[(nrow(ppdf_4)/2 + 1):nrow(ppdf_4)] <- 99
# # ppdf_4$sampling_day <- factor(ppdf_4$sampling_day, levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,"sum"))

# plot all points
pd4p <- ggplot(ppdf_4, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = sampling_day),
             color = colors[4],
             alpha = 0.4,width = 5) +
  # ggtitle("Participant 4") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd4p2 <- ggplot(ppdf_4, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[4],
             alpha = 0.4,width = 5) +
  ggtitle("All sampling days") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "IBM Plex Sans",
  size = 14,
  color = toRGB("grey50"))

(ly4 <- ggplotly(pd4p, width = 610, height = 672) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  animation_opts(800))

# ggsave(here("fig",file = "pd4p.png"), width = 12, height = 12)


```
</div>

```{r 4over, fig.width=6,fig.height=6, fig.align='center', fig.show='hold'}

(ly4_2 <- ggplotly(pd4p2, width = 600, height = 600) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  config(displayModeBar = FALSE))

# pd4p + ggtitle("All sampling days")

```
</div>

<h3 class = "pp-header">Participant 5</h3>
<div class ="pp-box" align = center>
<div class= "ani-plot-container">
```{r circ PP5, fig.width=6,fig.height=6, fig.align='center', fig.show='hold'}
# # add summary day of all points
# ppdf_5 <- rbind(ppdf_5,ppdf_5) %>% 
#   select(-c(X)) 
# 
# ppdf_5$sampling_day[(nrow(ppdf_5)/2 + 1):nrow(ppdf_5)] <- 99
# # ppdf_5$sampling_day <- factor(ppdf_5$sampling_day, levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,"sum"))

# plot all points
pd5p <- ggplot(ppdf_5, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") + 
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = sampling_day),
             color = colors[5],
             alpha = 0.4,width = 5) +
  # ggtitle("Participant 5") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd5p2 <- ggplot(ppdf_5, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[5],
             alpha = 0.4,width = 5) +
  ggtitle("All sampling days") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "IBM Plex Sans",
  size = 14,
  color = toRGB("grey50"))

(ly5 <- ggplotly(pd5p, width = 610, height = 672) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  animation_opts(800))

# ggsave(here("fig",file = "pd5p.png"), width = 12, height = 12)

```
</div>

```{r 5over, fig.width=6,fig.height=6, fig.align='center', fig.show='hold'}
(ly5_2 <- ggplotly(pd5p2, width = 600, height = 600) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  config(displayModeBar = FALSE))

# pd5p + ggtitle("All sampling days")

```
</div>

<h3 class = "pp-header">Participant 6</h3>
<div class ="pp-box" align = "center">
<div class= "ani-plot-container">
```{r r circ PP6, fig.width=6,fig.height=6, fig.align='center', fig.show='hold'}
# # add summary day of all points 
# ppdf_6 <- rbind(ppdf_6,ppdf_6) %>% 
#   select(-c(X))
# 
# ppdf_6$sampling_day[(nrow(ppdf_6)/2 + 1):nrow(ppdf_6)] <- 99
# # ppdf_6$sampling_day <- factor(ppdf_6$sampling_day, levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14, 15,"sum"))

pd6p <- ggplot(ppdf_6, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = sampling_day),
  # geom_jitter(aes(size = intensity),
             color = colors[6],
             alpha = 0.4,width = 5) +
  # ggtitle("Participant 6") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd6p2 <- ggplot(ppdf_6, aes(x = valence, y = arousal, label = emo_word)) +
  theme_minimal() +
  theme(text=element_text(family="IBM Plex Sans")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[6],
             alpha = 0.4,width = 5) +
  ggtitle("All sampling days") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "IBM Plex Sans",
  size = 14,
  color = toRGB("grey50"))

(ly6 <- ggplotly(pd6p, width = 610, height = 672) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  animation_opts(800))
 
# ggsave(here("fig",file = "pd6p.png"), width = 12, height = 12) 


```
</div>

```{r 6over, fig.width=6,fig.height=6, fig.align='center', fig.show='hold'}

(ly6_2 <- ggplotly(pd6p2, width = 600, height = 600) %>%
  layout(font = t) %>%
  config(displaylogo = FALSE) %>%
  config(displayModeBar = FALSE))

# pd6p + ggtitle("All sampling days")

```
</div>

<h3 class = "pp-header">All participants</h3>
<div class ="pp-box" align = "center">
<div class= "ani-plot-container">
```{r small multiples}
#--------------------#    
yaxis <- list(
  # title = 'arousal',
  # automargin = TRUE,
  # yanchor = "bottom"
  # tickangle = 90
  )
# 
# xaxis <- list(
#   title = 'valence',
#   automargin = TRUE,
#   xanchor = "right"
#   ) 

# title <- list(
#   text = "All participants",
#   xanchor = "left"
# )

subplot(ly1, ly2, ly3, ly4, ly5, ly6, 
        nrows = 2, 
        # margin = 0.025,
        titleX = F,
        titleY = F,
        shareX = T,
        shareY = T, 
        heights = c(0.40,0.40),
        widths = c(0.32,0.34,0.32)) %>% 
   layout(font = t,
          width = 800, height = 650,
          # yaxis = yaxis,
          # xaxis = xaxis,
          title = title,
          annotations = list(
            list(x = 0.5 , y = 0, text = "valence", showarrow = F, xref='paper', yref='paper', 
                 xanchor = "middle"),
            list(x = -0.07 , y = 0.5, text = "arousal", showarrow = F, xref='paper', yref='paper',
                 yanchor = "middle", textangle = 270))) %>%
   config(displaylogo = FALSE) %>%
   animation_opts(800) 
```
</div>
</div>

<div id = "footer">

<i>Completed for ARTG 6100 Spring 2021, IDDV Program, Northeastern University. Data
provided by the <a href = "https://www.affective-science.org/">Interdisciplinary Affective Science Lab</a></i>.

</div>

</div>