
---
title: Affective Trajectories
output: 
  html_document:
    css: style.css
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align = 'center')
```

<div id = "overall">

<div id = "opener" class = "text">
 
<h1 id = "title" class = "text">Affective Trajectories</h1>
<h4 id = "byline" class = "text">by Liz Cory</h4>

<p class = "text">
We all experience our feelings shifting — throughout the day and across periods of our lives. For some of us, the changes are slight, while others may feel totally different from one day to the next. To better understand how our feelings work and change, psychologists map them along two dimensions: "valence" and "arousal." Valence captures how pleasant you feel, and arousal marks how activated you feel. Together, valence and arousal are termed "affect."
</p>

<p class = "text">
This project documents how six individuals move through affective space over a two-week period, or, their <i>affective trajectories</i>. By animating daily reported feelings, we can explore where in affective space each person spent the most time. We can also see what emotion words a person uses to describe their feelings at each location by hovering over the dots.
</p>

</div>
```{r fonts, echo=FALSE,fig.show='hide'}
## Set up font
library(gfonts)

# setup_font(
#   id = "ibm-plex-sans",
#   output_dir = "./fonts"
# )

# setup_font(
#   id = "source-sans-pro",
#   output_dir = "./fonts"
# )

# setup_font(
#   id = "roboto",
#   output_dir = "./fonts"
# )

# setup_font(
#   id = "roboto-slab",
#   output_dir = "./fonts"
# )

use_font("roboto-slab", "fonts/css/roboto-slab.css")
use_font("roboto", "fonts/css/roboto.css")
# use_font("source-sans-pro", "fonts/css/source-sans-pro.css")
# use_font("ibm-plex-sans", "fonts/css/ibm-plex-sans.css")
```

```{r libs}
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(plotly)
library(wesanderson)
```

```{r load data, include=FALSE}
## identify list of files we want
pp_files <- list.files(here("dat"), pattern="*.csv", full.names = TRUE)

## read in all those files
df.list <- lapply(pp_files, read.csv)
names(df.list) <- paste0("ppdf_", seq_len(6))

## add each df to the global environments so we can more easily manipulate in ggplot
list2env(df.list, envir=.GlobalEnv)

# get colors
colors <- c(wes_palette("Darjeeling1"), wes_palette("Rushmore1")[4])

```
<hr>
<div class = "hdr-wpr text">
  <h1 class = "pp-header text">Participant 1</h1>
  <h4 class = "sub-head text">Daily reported feelings</h4>
</div>
<div class ="pp-box">
<div class= "ani-plot-container">
```{r circumplex PP1, out.width="100%"}
#--------------------#
# rename columns
ppdf_1 <- ppdf_1 %>% 
  rename("day" = "sampling_day",
         "emotion" = "emo_word")

# plot all points
pd1p <- ggplot(ppdf_1, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = day),
             color = colors[1],
             alpha = 0.4,width = 5) +
  # ggtitle("Daily reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd1p2 <- ggplot(ppdf_1, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[1],
             alpha = 0.4,width = 5) +
  # ggtitle("All reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "Roboto",
  size = 14,
  color = toRGB("grey50"))

# (ly1 <- ggplotly(pd1p, width = 610, height = 672) %>%
(ly1 <- ggplotly(pd1p) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent') %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  animation_opts(800))

# ggsave(here("fig",file = "pd1p.png"), width = 12, height = 12)

```
</div>

<div class = "hdr-wpr text">
  <h4 class = "sub-head text">All reported feelings</h4>
</div>
<div class= "full-plot-container">
```{r 1over, out.width="100%"}

# (ly1_2 <- ggplotly(pd1p2, width = 600, height = 600) %>%
(ly1_2 <- ggplotly(pd1p2) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent',
         xaxis = list(title = list(standoff = 20L)),
         yaxis = list(title = list(standoff = 20L))) %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  config(displayModeBar = FALSE))
  # widgetframe::frameWidget()

# frameWidget(ly1_2)

# pd1p + ggtitle("All reported feelings")

```
</div>

</div>

<div class = "hdr-wpr text">
  <h1 class = "pp-header text">Participant 2</h1>
  <h4 class = "sub-head text">Daily reported feelings</h4>
</div>
<div class ="pp-box">
<div class= "ani-plot-container">
```{r circ PP2, out.width='100%'}
# rename columns
ppdf_2 <- ppdf_2 %>% 
  rename("day" = "sampling_day",
         "emotion" = "emo_word")

# plot all points
pd2p <- ggplot(ppdf_2, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = day),
             color = colors[2],
             alpha = 0.4,width = 5) +
  # ggtitle("Daily reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd2p2 <- ggplot(ppdf_2, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[2],
             alpha = 0.4,width = 5) +
  # ggtitle("All reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "Roboto",
  size = 14,
  color = toRGB("grey50"))

# (ly2 <- ggplotly(pd2p, width = 610, height = 672) %>%
(ly2 <- ggplotly(pd2p) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent') %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  animation_opts(800))

# ggsave(here("fig",file = "pd2p.png"), width = 12, height = 12)


```
</div>

<div class = "hdr-wpr text">
  <h4 class = "sub-head text">All reported feelings</h4>
</div>
<div class= "full-plot-container">
```{r 2over, out.width='100%'}

# (ly2_2 <- ggplotly(pd2p2, width = 600, height = 600) %>%
(ly2_2 <- ggplotly(pd2p2) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent',
         xaxis = list(title = list(standoff = 20L)),
         yaxis = list(title = list(standoff = 20L))) %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  config(displayModeBar = FALSE))

# pd2p + ggtitle("All reported feelings")

```
</div>

</div>

<div class = "hdr-wpr text">
  <h1 class = "pp-header text">Participant 3</h1>
  <h4 class = "sub-head text">Daily reported feelings</h4>
</div>
<div class ="pp-box">
<div class= "ani-plot-container">
```{r circ PP3, out.width='100%'}
# rename columns
ppdf_3 <- ppdf_3 %>% 
  rename("day" = "sampling_day",
         "emotion" = "emo_word")

# plot all points
pd3p <- ggplot(ppdf_3, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = day),
             color = colors[3],
             alpha = 0.4,width = 5) +
  # ggtitle("Daily reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd3p2 <- ggplot(ppdf_3, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[3],
             alpha = 0.4,width = 5) +
  # ggtitle("All reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "Roboto",
  size = 14,
  color = toRGB("grey50"))

# (ly3 <- ggplotly(pd3p, width = 610, height = 672) %>%
(ly3 <- ggplotly(pd3p) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent') %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  animation_opts(800))

# ggsave(here("fig",file = "pd3p.png"), width = 12, height = 12)


```
</div>

<div class = "hdr-wpr text">
  <h4 class = "sub-head text">All reported feelings</h4>
</div>
<div class= "full-plot-container">
```{r 3over, out.width='100%'}

# (ly3_2 <- ggplotly(pd3p2, width = 600, height = 600) %>%
(ly3_2 <- ggplotly(pd3p2) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent',
         xaxis = list(title = list(standoff = 20L)),
         yaxis = list(title = list(standoff = 20L))) %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  config(displayModeBar = FALSE))

# pd3p + ggtitle("All reported feelings")

```
</div>

</div>

<div class = "hdr-wpr text">
  <h1 class = "pp-header text">Participant 4</h1>
  <h4 class = "sub-head text">Daily reported feelings</h4>
</div>
<div class ="pp-box">
<div class= "ani-plot-container">
```{r circ PP4, out.width="100%"}
# rename columns
ppdf_4 <- ppdf_4 %>% 
  rename("day" = "sampling_day",
         "emotion" = "emo_word")

# plot all points
pd4p <- ggplot(ppdf_4, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = day),
             color = colors[4],
             alpha = 0.4,width = 5) +
  # ggtitle("Daily reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd4p2 <- ggplot(ppdf_4, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[4],
             alpha = 0.4,width = 5) +
  # ggtitle("All reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "Roboto",
  size = 14,
  color = toRGB("grey50"))

# (ly4 <- ggplotly(pd4p, width = 610, height = 672) %>%
(ly4 <- ggplotly(pd4p) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent') %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  animation_opts(800))

# ggsave(here("fig",file = "pd4p.png"), width = 12, height = 12)


```
</div>

<div class = "hdr-wpr text">
  <h4 class = "sub-head text">All reported feelings</h4>
</div>
<div class= "full-plot-container">
```{r 4over, out.width='100%'}

# (ly4_2 <- ggplotly(pd4p2, width = 600, height = 600) %>%
(ly4_2 <- ggplotly(pd4p2) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent',
         xaxis = list(title = list(standoff = 20L)),
         yaxis = list(title = list(standoff = 20L))) %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  config(displayModeBar = FALSE))

# pd4p + ggtitle("All reported feelings")

```
</div>

</div>

<div class = "hdr-wpr text">
  <h1 class = "pp-header text">Participant 5</h1>
  <h4 class = "sub-head text">Daily reported feelings</h4>
</div>
<div class ="pp-box">
<div class= "ani-plot-container">
```{r circ PP5, out.width="100%"}
# rename columns
ppdf_5 <- ppdf_5 %>% 
  rename("day" = "sampling_day",
         "emotion" = "emo_word")

# plot all points
pd5p <- ggplot(ppdf_5, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") + 
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = day),
             color = colors[5],
             alpha = 0.4,width = 5) +
  # ggtitle("Daily reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd5p2 <- ggplot(ppdf_5, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[5],
             alpha = 0.4,width = 5) +
  # ggtitle("All reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "Roboto",
  size = 14,
  color = toRGB("grey50"))

# (ly5 <- ggplotly(pd5p, width = 610, height = 672) %>%
(ly5 <- ggplotly(pd5p) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent') %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  animation_opts(800))

# ggsave(here("fig",file = "pd5p.png"), width = 12, height = 12)

```
</div>

<div class = "hdr-wpr text">
  <h4 class = "sub-head text">All reported feelings</h4>
</div>
<div class= "full-plot-container">
```{r 5over, out.width='100%'}

# (ly5_2 <- ggplotly(pd5p2, width = 600, height = 600) %>%
(ly5_2 <- ggplotly(pd5p2) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent',
         xaxis = list(title = list(standoff = 20L)),
         yaxis = list(title = list(standoff = 20L))) %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  config(displayModeBar = FALSE))

# pd5p + ggtitle("All reported feelings")

```
</div>

</div>

<div class = "hdr-wpr text">
  <h1 class = "pp-header text">Participant 6</h1>
  <h4 class = "sub-head text">Daily reported feelings</h4>
</div>
<div class ="pp-box">
<div class= "ani-plot-container">
```{r r circ PP6, out.width='100%'}
# # add summary day of all points 
# ppdf_6 <- rbind(ppdf_6,ppdf_6) %>% 
#   select(-c(X))
# 
# ppdf_6$day[(nrow(ppdf_6)/2 + 1):nrow(ppdf_6)] <- 99
# # ppdf_6$day <- factor(ppdf_6$day, levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14, 15,"sum"))

# rename columns
ppdf_6 <- ppdf_6 %>% 
  rename("day" = "sampling_day",
         "emotion" = "emo_word")

pd6p <- ggplot(ppdf_6, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity, frame = day),
  # geom_jitter(aes(size = intensity),
             color = colors[6],
             alpha = 0.4,width = 5) +
  # ggtitle("Daily reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

# plot all points
pd6p2 <- ggplot(ppdf_6, aes(x = valence, y = arousal, label = emotion)) +
  theme_minimal() +
  theme(text=element_text(family="Roboto")) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = 0, color = "darkgray") +
  geom_jitter(aes(size = intensity),
             color = colors[6],
             alpha = 0.4,width = 5) +
  # ggtitle("All reported feelings") +
  xlim(-50,50) +
  ylim(-50,50)

t <- list(
  family = "Roboto",
  size = 14,
  color = toRGB("grey50"))

# (ly6 <- ggplotly(pd6p, width = 610, height = 672) %>%
(ly6 <- ggplotly(pd6p) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent') %>% 
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  animation_opts(800))
 
# ggsave(here("fig",file = "pd6p.png"), width = 12, height = 12) 


```
</div>

<div class = "hdr-wpr text">
  <h4 class = "sub-head text">All reported feelings</h4>
</div>
<div class= "full-plot-container">
```{r 6over, out.width='100%'}

# (ly6_2 <- ggplotly(pd6p2, width = 600, height = 600) %>%
(ly6_2 <- ggplotly(pd6p2) %>%
  layout(font = t,
         plot_bgcolor ='transparent',
         paper_bgcolor='transparent',
         xaxis = list(title = list(standoff = 20L)),
         yaxis = list(title = list(standoff = 20L)),
         title = list(list(xref = "container", xanchor = "left", x = -.1))) %>%
  config(displaylogo = FALSE) %>%
  config(responsive = TRUE) %>% 
  config(displayModeBar = FALSE))

# pd6p + ggtitle("All reported feelings")

```
</div>

</div>

<div class = "hdr-wpr text">
  <h1 class = "pp-header text">All participants</h1>
  <h4 class = "sub-head text">Daily reported feelings</h4>
</div>
<div class ="pp-box">
<div class= "ani-plot-container">
```{r small multiples, out.width="100%"}
#--------------------#    
subplot(ly1, ly2, ly3, ly4, ly5, ly6, 
        nrows = 2, 
        # margin = 0.025,
        titleX = F,
        titleY = F,
        shareX = T,
        shareY = T, 
        heights = c(0.36,0.36),
        widths = c(0.325,0.325,0.325)) %>% 
   layout(font = t,
          # width = 800, height = 650,
          # yaxis = yaxis,
          # xaxis = xaxis,
          title = title,
          annotations = list(
            list(x = 0.5 , y = -0.03, text = "valence", 
                 showarrow = F, xref='paper', yref='paper', 
                 xanchor = "middle", font = list(color ="black")),
            list(x = -0.13 , y = 0.5, text = "arousal", 
                 showarrow = F, xref='paper', yref='paper',
                 yanchor = "middle", textangle = 270, 
                 font = list(color ="black")))) %>%
   config(displaylogo = FALSE) %>%
   config(responsive = TRUE) %>% 
   animation_opts(800) 
```
</div>
</div>
<hr>
<div id = "footer" class = "text">

Completed for ARTG 6100 Spring 2021, IDDV Program, Northeastern University. Data was
provided by the <a href = "https://www.affective-science.org/">Interdisciplinary Affective Science Lab</a>.

</div>
</div>